Effects of pre-existing structures on the seismicity of the Charlevoix Seismic Zone

### 2019-06-28 main_full_manuscript.tex
Some edits by Chris

\begin{figure}[h]
\centering
\includegraphics[width=18.5pc]{Figures/Map_of_CSZ_2.png} 
\caption{Topography\add[OF]{, bathymetry,} and seismicity of the Charlevoix Seismic Zone (CSZ) as well as the locations of the impact \change[OF]{crater}{structure} (outer cyan circle) and the more damaged inner \change[OF]{crater}{impact structure} (inner cyan circle). Small circles \add[OF]{and diamonds} are the relocated epicenters~\citep{Powell_2017} \add[OF]{and the complete dataset for the years 1988-2011, respectively,} and their colors represent the focal depths. The focal mechanisms are for the earthquakes used by \citet{Mazzotti_2010} for the stress inversion. Large circles labeled as NW and SE show orientations of SH$_{\max}$ from the stress inversion of focal mechanism (SH$_S$) and from borehole breakout measurements (SH$_B$) for the earthquake clusters northwest and southeast of the \change[OF]{crater}{impact structure} center. Solid black lines mark the rift faults known in the region: GNF, Gouffre Northwest Fault; SLF, Saint-Laurent fault; CHF, Charlevoix Fault; and SSF, South Shore Fault \citep{Rondot_1971,lamontagne1999}. The inset shows the location of the CSZ in eastern Canada. Earthquake \change[OF]{epicenters}{hypocenters} from the National Resources Canada catalog for the years 1988-2011.}
\label{figone}
\end{figure}

\citet{Vlahovic_2003} conducted a three-dimensional (3D) P-wave velocity study of the CSZ. The tomography model shows several \change[OF]{high-velocity}{higher P velocity} bodies at mid-crustal depths. A significant fraction of the CSZ earthquakes occurs around these \change[OF]{high-velocity}{higher P velocity} bodies and some larger earthquakes occur along the northern edges of the high velocity bodies.

@@

We conducted resolution tests in which the change in displacement and stress solutions relative to a 1 km-resolution model are monitored as a function of element size. \change[OF]{Displacement was almost constant for 2 km or smaller element sizes}{Displacements and differential stresses do not change much for elements sizes equal to 4 km or smaller (Fig. S1)}. \add[OF]{Based on the resolution tests, we use the element size of 2 km for most of the models and 4 km for more computationally expensive ones like those for exploring the effects of friction coefficient.} We also observe edge effects by comparing a 420 $\times$ 420 $\times$ 40 km model with the original 220 $\times$ 220 $\times$ 40 km model, finding that the seismogenic parts around the \change[OF]{crater}{impact structure} show consistent displacement and stress solutions for both domain sizes. Based on this result, we choose the smaller domain for this study and set the element size to be about 2 km. \add[OF]{The average computation time was about 16 days for fault models ($\mu$ = 0.3) with 2 km element size on 16 cores on the University of Memphis High Performance Computing cluster. The model with fault dips, 65$^\circ$-40$^\circ$-40$^\circ$ took about twice longer to run for the same model time. We do not clearly understand why the model with shallower-dipping faults is slower than the others roughly by a factor of two and further analyzing computational efficiency is beyond the scope of this paper.}
%we use an element size of 4 km for models with friction coefficient ($\mu$) values of 0.1 and 0.2 due to their expensive computation time.

@@

Velocities are prescribed on the sides perpendicular to the $x$-axis (Fig. \ref{figtwo}A) while the $y$-perpendicular sides and the bottom boundary are free-slip and the top boundary is traction-free. Rather than applying velocities oblique to the sides, we rotate the model domain such that the $x$-axis is oriented to N55$^\circ$E, the regional SH$_{\max}$ direction~\citep{Zoback_1992} (Fig. \ref{figtwo}B). We apply a 1 m/yr compressive velocity in the x-axis direction. The magnitude of boundary velocity does not represent any tectonic loading in the continental interiors but has the sole purpose of increasing the differential stress ($\sigma_D$ = $\sigma_1 - \sigma_3$). Noting that \change[OF]{most earthquakes in the crater occur near 10 km depth}{the distribution of earthquakes within the impact structure has a peak at 10 km} \citep{Baird_2010,Powell_2017}, we increase the boundary displacement until $\sigma_D$ reaches 706 MPa at 10 km depth based on Byerlee's law for dry rocks, $\tau$  = 50 MPa + 0.6$\,\sigma_n$~\citep{Byerlee_1978}. \add[OF]{Supplementary Information provides more details of the full derivation of the stopping criterion for boundary loading. We assume that} $\sigma_3$ corresponds to the lithostatic stress \change[OF]{. C}{and adopt the sign convention that compressive stress is negative.}

@@

For simplicity, we assume that the \change[OF]{crater}{impact structure}'s bulk and shear moduli are a certain fraction of those of the crust~\citep[e.g.,][]{Baird_2010}. Similarly, the density of the \change[OF]{crater}{impact structure} is reduced \remove[OF]{by 10 \%} from that of the 1D velocity models over the corresponding depth range \add[OF]{to model a damaged crustal rock. Models with density reduced by 0 to 20 \% were run but did not exhibit significant changes as shown in Fig. S3 in the supplementary information. Unless stated otherwise, models presented in this study has 10 \% of density reduction in the impact structure.}

@@

The reference model (SNFR25) has the geometry described above but without the rift faults. This is equivalent to a model with infinitely strong faults. The initial stress state is close to the initially assumed lithostatic stress after the nonplanar correction due to the \change[OF]{crater}{impact structure}. We determine the percentage change in differential stress, $\Delta\sigma_{D}^{\prime}=100\,( \sigma_{D}-\sigma_{D,ref}) / \sigma_{D,ref}$, and the change in the orientation of SH$_{max}$,  $\Delta\phi_{\max}=\phi_{\max}-\phi_{\max,ref}$, where $\phi_{\max}$ denotes the orientation of the maximum principal stress ($\sigma_1$). Since our model domain was rotated clockwise about the center of the \change[OF]{crater}{impact structure} by 35$^\circ$, $\phi_{\max}$ equal to zero should be understood as parallel to the regional stress field. Positive and negative values of $\phi_{\max}$ correspond to anticlockwise and clockwise rotation from the direction of applied loading, respectively. \add[OF]{In order to quantify the spatial correlation between the simulated stress changes and the earthquake hypocenters, we determine the percentage of hypocenters that fall within region of $\Delta\sigma_{D}^{\prime} \ge$ 0.5 \%. We did not account for the uncertainty in hypocenter relocation and the effect of any along-strike variation in the fault dips in reality.}

@@

\begin{figure}[ht]
\centering
%\includegraphics[width=25pc]{Figures/SD70R25.png}
\includegraphics[width=25pc]{Figures/SD70R25_2.png}
\caption{(A) Differential stress ($\sigma_D$), (B) SH$_{\max}$ orientation ($\phi_{\max}$), (C) Percentage change in differential stress  ($\Delta\sigma_{D}^{\prime}$) and (D) the change in $\phi_{\max}$ ($\Delta\phi_{\max}$) in SD70R25 model relative to SNFR25 at various depths. Note the change in scale in \ref{fig:SD70R25}A for depths greater than 10 km. Earthquakes within 2 km of each depth slice are represented by black dots in \ref{fig:SD70R25}C. \add[OF]{The outline of the impact structure is represented as black circle.} Each figure is 70 $\times$ 70 km.} 
\label{fig:SD70R25}
\end{figure}

@@

The distribution of positive $\Delta\sigma_{D}^{\prime}$ values, corresponding to increased differential stresses in SD70R25 relative to SNFR25, spatially overlaps with the seismicity in general and particularly well at 10 km (Fig. \ref{fig:SD70R25}C). \add[OF]{About 70\% of the hypocenters fall in region with $\Delta\sigma_{D}^{\prime} \ge$ 0.5\%.} We superimpose the relocated hypocenters of earthquakes of \citet{Powell_2017} that occurred within 2 km of each depth slice. The hypocenters are rotated clockwise by 35$^{\circ}$ about the center of the \change[OF]{crater}{impact structure} in order for them to be consistent with the rotated model geometry. Some earthquakes fall in the region of negative $\Delta\sigma_{D}^{\prime}$ near the SW and NE boundary of the \change[OF]{crater}{impact structure} at depths shallower than 15 km.

@@

\begin{figure}[ht]
\centering
%\includegraphics[width=30pc]{Figures/SD65R25.png}
\includegraphics[width=30pc]{Figures/SD65R25_2.png}
\caption{Percentage change in differential stress ($\Delta\sigma_{D}^{\prime}$), SH$_{\max}$ orientation ($\phi_{\max})$ and the change in $\phi_{\max}$ ($\Delta\phi_{\max}$) at 10 and 15 km depths from the SD65R25 model. Earthquakes within 2 km of each depth slice are represented by black dots in \ref{fig:eff_of_dips}A. \add[OF]{The outline of the impact structure is represented as black circle.} Each figure is 70 $\times$ 70 km.}
\label{fig:eff_of_dips}
\end{figure}

@@

The region of increased differential stress in SD65R25 exhibits well-defined linear trends near the rift faults, and thus gives a better explanation for the narrow and well-defined seismicity near the \change[OF]{crater}{impact structure} (Fig. \ref{fig:profiles_65_40_40_2km}). \add[OF]{About 25\% of the hypocenters fall in region with $\Delta\sigma_{D}^{\prime} \ge$ 0.5\%.}. The SD65R25 model shows a region of negative $\Delta\sigma_{D}^{\prime}$ between the observed seismicity on the northernmost and southernmost faults. The earthquakes on the middle rift fault and most of the earthquakes on the southernmost fault outside the \change[OF]{crater}{impact structure} fall in regions with negative $\Delta\sigma_{D}^{\prime}$ (Fig. \ref{fig:profiles_65_40_40_2km}).

@@

\subsubsection{SD70R50}
While the effects of the \change[OF]{crater}{impact structure} and rift faults are still clear and consistent, SD70R50 values of $\Delta \sigma_{D}^{\prime}$ are mostly negative with some exceptions at 10 km depth amounting to about a 10 \% increased stress (Fig. \ref{fig:eff_of_moduli_ratio}A). The region of positive $\Delta\sigma_{D}^{\prime}$ in the SD70R50 model overlaps with the hypocenters only in the southeast corner of the \change[OF]{crater}{impact structure} boundary and the southernmost rift fault (i.e., the Charlevoix fault) at 5 km depth \add[OF]{, amounting to about 45\% of the hypocenters in region with $\Delta\sigma_{D}^{\prime} \ge$ 0.5\%.} (Fig. \ref{fig:eff_of_moduli_ratio}A). The spatial correlation of seismicity with increased $\sigma_{D}$ relative to the reference model is stronger at 10 km depth.

@@

Positive $\Delta\sigma_{D}^{\prime}$ values in SD70R100 are seen only at a depth of 10 km and their distribution does not fit the observed seismicity (Fig. \ref{fig:eff_of_moduli_ratio}B).  \add[OF]{About 23\% of the hypocenters fall in region with $\Delta\sigma_{D}^{\prime} \ge$ 0.5\%.} $\Delta\sigma_{D}^{\prime}$ correlates with observed seismicity below 15 km except for some earthquakes in the northern cluster. However, the model predicts earthquakes throughout the entire length of the rift faults at and below 15 km depth. Figures that shows all the depth slices are in the supplementary materials (Fig. S4). \add[OF]{The result of a model with an intermediate moduli ratio of 0.75 is somewhat between the SD70R50 and SD70R100 models (Fig. S5).}

@@

Magnitudes of $\Delta \sigma_{D}^{\prime}$ decrease as $\mu$ increases (Fig. \ref{fig:eff_of_mu}A) because the model behaviors must approach those of the no-fault model as fault strength increases. When $\mu$ values are 0.1 and 0.2, the area of positive $\Delta\sigma_{D}^{\prime}$ regions does not cover the entire observed seismicity\add[OF]{, with about 3\% and 24\% of the hypocenters in region with $\Delta\sigma_{D}^{\prime} \ge$ 0.5\%, respectively}. The regions of positive $\Delta\sigma_{D}^{\prime}$ for $\mu$ values equal to 0.3 and 0.4 show better spatial correspondence with the seismicity of the CSZ. However, the model with $\mu = 0.4$ predicts positive Coulomb stress changes along the entire length of the rift faults, which is contrary to the limited spatial extent of the observed seismicity. \add[OF]{Also, about 66\% of the hypocenters in the model with $\mu$ of 0.3 fall within region of $\Delta\sigma_{D}^{\prime} \ge$ 0.5\%, compare to a value of 35\% when $\mu$ is 0.4. However, when we consider region with $\Delta\sigma_{D}^{\prime} \ge$ 0\%, the percentage increases to 82\% when $\mu$ of 0.4 compare to 73\% in the model with $\mu$ of 0.3.} $\Delta\sigma_{D}^{\prime}$ approaches zero when $\mu$ values are 0.5 and 0.6, for which the faults are locked. \add[OF]{The percentage of hypocenters in region with $\Delta\sigma_{D}^{\prime} \ge$ 0.5\% reduces to 1.4\% and 0\% when $\mu$ values are 0.5 and 0.6, respectively.}

\begin{figure}[ht]
\centering
%\includegraphics[width=30pc]{Figures/SD70R25V.png}
\includegraphics[width=30pc]{Figures/SD70R25V_2.png}
\caption{Effect of friction coefficient on the $\Delta\sigma_{D}^{\prime}$ and $\phi_{\max}$ in the SD70R25V model at 10 km depth. Note the change in scale for $\mu$ greater than 0.3. Earthquakes within 2 km of each depth slice are represented by black dots in \ref{fig:eff_of_mu}A. \add[OF]{The outline of the impact structure is represented as black circle.} Each figure is 70 $\times$ 70 km.} 
\label{fig:eff_of_mu}
\end{figure}

@@

Our present models, however, cannot explain the differences in earthquake activity between the \change[OF]{crater}{impact structure} and the surrounding crust in the CSZ. `b' values of the Gutenberg-Richter relationship for the regions within and outside the \change[OF]{crater}{impact structure} are 0.93 and 0.74, respectively~\citep{Yu_2016}, indicating a higher frequency of small magnitude earthquakes within the \change[OF]{crater}{impact structure}. Also, \change[OF]{the large-magnitude}{most M4+} earthquakes in the CSZ \change[OF]{occurred}{appear to occur} in the northeast region outside the \change[OF]{crater}{impact structure}~\citep[e.g.,][]{lamontagne1999,Powell_2017,Baird_2010}. \add[OF]{Reexamination of historical earthquakes in the Charlevoix area by}~\citet{Stevens1980} \add[OF]{shows that some M4+ earthquakes of the period 1924-1978 also occurred at the SE end of the impact structure.} Our models do not show significantly higher values of $\Delta\sigma_{D}^\prime$ in the seismogenic zone outside the \change[OF]{crater}{impact structure} relative to the regions within the impact \change[OF]{crater}{structure} (Fig. \ref{fig:SD70R25}C). A possible explanation is that faults created in the highly fractured \change[OF]{crater}{impact structure} region~\citep{RONDOT_2000} might have lower friction coefficients than the faults outside the \change[OF]{crater}{impact structure}. Stress build-up on those weaker \change[OF]{crater}{impact structure} faults can be quicker and smaller in magnitude than on faults outside the \change[OF]{crater}{impact structure} under the same tectonic loading, resulting in more numerous, smaller magnitude earthquakes within versus outside the \change[OF]{crater}{impact structure}.

@@

\subsection{Frictional strengths of the rift faults}
The surface expression of the rift faults used in this study correspond to the Gouffre Northwest Fault, Saint-Laurent fault and Charlevoix Fault (Fig. 1). We did not consider the South Shore fault given its aseismic nature and \change[OF]{its location at the extreme eastern edge of the crater}{it appears to mark a boundary to the active zone}~\citep{lamontagne1999}. Our models show that a $\mu$ of 0.3 and cohesion values of 0 MPa create $\Delta\sigma_{D}^\prime$ maps that spatially correlate with the seismicity of CSZ and the observed stress rotations (Fig. \ref{fig:eff_of_mu}). \add[OF]{We ran a series of models in which the rift faults dip at 70$^{\circ}$ and the friction coefficient ($\mu$) is fixed to be 0.3 while cohesion ($C$) is one of 0, 3, 5, 10, 20 and 30 MPa (Fig. S6). Cohesion values less than 5 MPa could explain the spatial distribution of observed seismicity in the CSZ.} \add[OF]{A $\mu$ of 0.4 could also be suitable for the rift faults as the $\Delta\sigma_{D}^\prime$ maps also spatially correlate with the seismicity of CSZ, and some earthquakes also occurred at the SW region outside the impact structure (Fig. $\ref{figone}$). However, previous works in the CSZ agreed with a finite extent of the seismic zone} ~\citep[e.g.,][]{Stevens1980}. \change[OF]{This}{Our fault model} is different from the work of \citet{Baird_2010} who modeled the Gouffre Northwest Fault, Saint-Laurent fault and South Shore Fault (Fig. \ref{figone}).

@@

\subsection{Velocity model of the CSZ}
Our models show that a 1D velocity model of the CSZ \citep{lamontagne1999} (Fig. \ref{figthree}) with embedded \change[OF]{crater}{impact structure} and rift faults can explain the observed seismicity and more than 50$\%$ of the stress rotations (Fig. \ref{fig:eff_of_vel}A). Using the velocity model of the \change[OF]{Sanguenay}{Saguenay} region does not improve this result (Fig. \ref{fig:SD70R25}). Similarly, $\Delta\sigma_{D}^\prime$ in the TD70 model did not spatially fit the recorded earthquakes in the upper 10 km in the CSZ because the distinct and lower-strength damaged zone is not apparent in the 3D tomography \citep{Powell_2017} (Fig. \ref{fig:eff_of_vel}B). \citet{lamontagne1999} \change[OF]{Further investigation is needed to determine why the supposedly better velocity models fail}{is a 1-D velocity model for the Canadian Shield region that does not include lateral variations of} velocity model to explain the seismicity and stress orientations.

@@

\item The combined effects of the impact \change[OF]{crater}{impact structure} and the rift faults are required in the CSZ to explain the observed distribution of seismicity and more than 50$\%$ of the observed clockwise rotations of the $\phi_{\max}$ in the CSZ \add[OF]{in respect to the regional maximum compressive stress direction}.

>>>>

\begin{figure}[h]
\centering
\includegraphics[width=18.5pc]{Figures/Map_of_CSZ_2.png} 
\caption{Topography\add[OF]{, bathymetry,} and seismicity of the Charlevoix Seismic Zone (CSZ) as well as the locations of the impact \change[OF]{crater}{structure} (outer cyan circle) and the more damaged inner \change[OF]{crater}{impact structure} (inner cyan circle). Small circles \add[OF]{and diamonds} are the relocated epicenters~\citep{Powell_2017} \add[OF]{and the complete dataset for the years 1988-2011, respectively,} and their colors represent the focal depths. The focal mechanisms are for the earthquakes used by \citet{Mazzotti_2010} for the stress inversion. Large circles labeled as NW and SE show orientations of SH$_{\max}$ from the stress inversion of focal mechanism (SH$_S$) and from borehole breakout measurements (SH$_B$) for the earthquake clusters northwest and southeast of the \change[OF]{crater}{impact structure} center. Solid black lines mark the rift faults known in the region: GNF, Gouffre Northwest Fault; SLF, Saint-Laurent fault; CHF, Charlevoix Fault; and SSF, South Shore Fault \citep{Rondot_1971,lamontagne1999}. The inset shows the location of the CSZ in eastern Canada. Earthquake \change[OF]{epicenters}{dataset} from the National Resources Canada catalog for the years 1988-2011.}
\label{figone}
\end{figure}

\citet{Vlahovic_2003} conducted a three-dimensional (3D) P-wave velocity study of the CSZ. The tomography model shows several \change[OF]{high-velocity}{higher P-wave velocity} bodies at mid-crustal depths. A significant fraction of the CSZ earthquakes occurs around these \change[OF]{high-velocity}{higher P-wave velocity} bodies and some larger earthquakes occur along the northern edges of the high velocity bodies.

@@

We conducted resolution tests in which the change in displacement and stress solutions relative to a 1 km-resolution model are monitored as a function of element size. \change[OF]{Displacement was almost constant for 2 km or smaller element sizes}{Displacements and differential stresses do not change appreciably for elements sizes equal to 4 km or smaller (Fig. S1)}. \add[OF]{Based on the resolution tests, we use the element size of 2 km for most of the models and 4 km for more computationally expensive models like those for exploring the effects of the friction coefficient.} We also observe edge effects by comparing a 420 $\times$ 420 $\times$ 40 km model with the original 220 $\times$ 220 $\times$ 40 km model, finding that the seismogenic parts around the \change[OF]{crater}{impact structure} show consistent displacement and stress solutions for both domain sizes. Based on this result, we choose the smaller domain for this study and set the element size to be about 2 km. \add[OF]{The average computation time was about 16 days for fault models ($\mu$ = 0.3) with 2 km element size on 16 cores on the University of Memphis High Performance Computing cluster. The model with fault dips of 65$^\circ$-40$^\circ$-40$^\circ$ took about twice longer to run for the same model time. We do not clearly understand why the model with shallower-dipping faults is slower than the others roughly by a factor of two and further analyzing computational efficiency is beyond the scope of this paper.}
%we use an element size of 4 km for models with friction coefficient ($\mu$) values of 0.1 and 0.2 due to their expensive computation time.

@@

Velocities are prescribed on the sides perpendicular to the $x$-axis (Fig. \ref{figtwo}A) while the $y$-perpendicular sides and the bottom boundary are free-slip and the top boundary is traction-free. Rather than applying velocities oblique to the sides, we rotate the model domain such that the $x$-axis is oriented to N55$^\circ$E, the regional SH$_{\max}$ direction~\citep{Zoback_1992} (Fig. \ref{figtwo}B). We apply a 1 m/yr compressive velocity in the x-axis direction. The magnitude of boundary velocity does not represent any tectonic loading in the continental interiors but has the sole purpose of increasing the differential stress ($\sigma_D$ = $\sigma_1 - \sigma_3$). Noting that \change[OF]{most earthquakes in the crater occur near 10 km depth}{the depth distribution of earthquakes within the impact structure has a peak at 10 km} \citep{Baird_2010,Powell_2017}, we increase the boundary displacement until $\sigma_D$ reaches 706 MPa at 10 km depth based on Byerlee's law for dry rocks, $\tau$  = 50 MPa + 0.6$\,\sigma_n$~\citep{Byerlee_1978}. \add[OF]{More details of the full derivation of the stopping criterion for boundary loading is given in the supplementary information. We assume that} $\sigma_3$ corresponds to the lithostatic stress \change[OF]{. C}{and adopt the sign convention that compressive stress is negative.}

@@

For simplicity, we assume that the \change[OF]{crater}{impact structure}'s bulk and shear moduli are a certain fraction of those of the crust~\citep[e.g.,][]{Baird_2010}. Similarly, the density of the \change[OF]{crater}{impact structure} is reduced \remove[OF]{by 10 \%} from that of the 1D velocity models over the corresponding depth range \add[OF]{to model a damaged crustal rock. Models with density reduced by 0 to 20 \% were run but did not exhibit significant changes as shown in Fig. S3 in the supplementary information. Unless stated otherwise, models presented in this study have 10 \% of density reduction in the impact structure.}

@@

The reference model (SNFR25) has the geometry described above but without the rift faults. This is equivalent to a model with infinitely strong faults. The initial stress state is close to the initially assumed lithostatic stress after the nonplanar correction due to the \change[OF]{crater}{impact structure}. We determine the percentage change in differential stress, $\Delta\sigma_{D}^{\prime}=100\,( \sigma_{D}-\sigma_{D,ref}) / \sigma_{D,ref}$, and the change in the orientation of SH$_{max}$,  $\Delta\phi_{\max}=\phi_{\max}-\phi_{\max,ref}$, where $\phi_{\max}$ denotes the orientation of the maximum principal stress ($\sigma_1$). Since our model domain was rotated clockwise about the center of the \change[OF]{crater}{impact structure} by 35$^\circ$, $\phi_{\max}$ equal to zero should be understood as parallel to the regional stress field. Positive and negative values of $\phi_{\max}$ correspond to anticlockwise and clockwise rotation from the direction of applied loading, respectively. \add[OF]{In order to quantify the spatial correlation between the simulated stress changes and the earthquake hypocenters, we determine the percentage of hypocenters that fall within the region of $\Delta\sigma_{D}^{\prime} \ge$ 0.5 \%. We did not account for the uncertainty in hypocenter relocation and the effect of any along-strike variation in the fault dips.} % in reality

@@

\begin{figure}[ht]
\centering
%\includegraphics[width=25pc]{Figures/SD70R25.png}
\includegraphics[width=25pc]{Figures/SD70R25_2.png}
\caption{(A) Differential stress ($\sigma_D$), (B) SH$_{\max}$ orientation ($\phi_{\max}$), (C) Percentage change in differential stress  ($\Delta\sigma_{D}^{\prime}$) and (D) the change in $\phi_{\max}$ ($\Delta\phi_{\max}$) in SD70R25 model relative to SNFR25 at various depths. Note the change in scale in \ref{fig:SD70R25}A for depths greater than 10 km. Earthquakes within 2 km of each depth slice are represented by black dots in \ref{fig:SD70R25}C. \add[OF]{The outline of the impact structure is represented as a black circle.} Each figure is 70 $\times$ 70 km.} 
\label{fig:SD70R25}
\end{figure}

@@

The distribution of positive $\Delta\sigma_{D}^{\prime}$ values, corresponding to increased differential stresses in SD70R25 relative to SNFR25, spatially overlaps with the seismicity in general and particularly well at 10 km (Fig. \ref{fig:SD70R25}C). \add[OF]{About 70\% of the hypocenters fall in the region with $\Delta\sigma_{D}^{\prime} \ge$ 0.5\%.} We superimpose the relocated hypocenters of earthquakes of \citet{Powell_2017} that occurred within 2 km of each depth slice. The hypocenters are rotated clockwise by 35$^{\circ}$ about the center of the \change[OF]{crater}{impact structure} in order for them to be consistent with the rotated model geometry. Some earthquakes fall in the region of negative $\Delta\sigma_{D}^{\prime}$ near the SW and NE boundary of the \change[OF]{crater}{impact structure} at depths shallower than 15 km.

@@

\begin{figure}[ht]
\centering
%\includegraphics[width=30pc]{Figures/SD65R25.png}
\includegraphics[width=30pc]{Figures/SD65R25_2.png}
\caption{Percentage change in differential stress ($\Delta\sigma_{D}^{\prime}$), SH$_{\max}$ orientation ($\phi_{\max})$ and the change in $\phi_{\max}$ ($\Delta\phi_{\max}$) at 10 and 15 km depths from the SD65R25 model. Earthquakes within 2 km of each depth slice are represented by black dots in \ref{fig:eff_of_dips}A. \add[OF]{The outline of the impact structure is represented as a black circle.} Each figure is 70 $\times$ 70 km.}
\label{fig:eff_of_dips}
\end{figure}

@@

The region of increased differential stress in SD65R25 exhibits well-defined linear trends near the rift faults, and thus gives a better explanation for the narrow and well-defined seismicity near the \change[OF]{crater}{impact structure} (Fig. \ref{fig:profiles_65_40_40_2km}). \add[OF]{About 25\% of the hypocenters fall in the region with $\Delta\sigma_{D}^{\prime} \ge$ 0.5\%.}. The SD65R25 model shows a region of negative $\Delta\sigma_{D}^{\prime}$ between the observed seismicity on the northernmost and southernmost faults. The earthquakes on the middle rift fault and most of the earthquakes on the southernmost fault outside the \change[OF]{crater}{impact structure} fall in regions with negative $\Delta\sigma_{D}^{\prime}$ (Fig. \ref{fig:profiles_65_40_40_2km}).

@@

\subsubsection{SD70R50}
While the effects of the \change[OF]{crater}{impact structure} and rift faults are still clear and consistent, SD70R50 values of $\Delta \sigma_{D}^{\prime}$ are mostly negative with some exceptions at 10 km depth amounting to about a 10 \% increased stress (Fig. \ref{fig:eff_of_moduli_ratio}A). The region of positive $\Delta\sigma_{D}^{\prime}$ in the SD70R50 model overlaps with the hypocenters only in the southeast corner of the \change[OF]{crater}{impact structure} boundary and the southernmost rift fault (i.e., the Charlevoix fault) at 5 km depth \add[OF]{, amounting to about 45\% of the hypocenters in the region with $\Delta\sigma_{D}^{\prime} \ge$ 0.5\%.} (Fig. \ref{fig:eff_of_moduli_ratio}A). The spatial correlation of seismicity with increased $\sigma_{D}$ relative to the reference model is stronger at 10 km depth.

@@

Positive $\Delta\sigma_{D}^{\prime}$ values in SD70R100 are seen only at a depth of 10 km and their distribution does not fit the observed seismicity (Fig. \ref{fig:eff_of_moduli_ratio}B).  \add[OF]{About 23\% of the hypocenters fall in the region with $\Delta\sigma_{D}^{\prime} \ge$ 0.5\%.} $\Delta\sigma_{D}^{\prime}$ correlates with observed seismicity below 15 km except for some earthquakes in the northern cluster. However, the model predicts earthquakes throughout the entire length of the rift faults at and below 15 km depth. Figures that shows all the depth slices are in the supplementary materials (Fig. S4). \add[OF]{The result of a model with an intermediate moduli ratio of 0.75 is somewhat between the SD70R50 and SD70R100 models (Fig. S5).}

@@

Magnitudes of $\Delta \sigma_{D}^{\prime}$ decrease as $\mu$ increases (Fig. \ref{fig:eff_of_mu}A) because the model behaviors must approach those of the no-fault model as fault strength increases. When $\mu$ values are 0.1 and 0.2, the area of positive $\Delta\sigma_{D}^{\prime}$ regions does not cover the entire observed seismicity\add[OF]{, with about 3\% and 24\% of the hypocenters in the region with $\Delta\sigma_{D}^{\prime} \ge$ 0.5\%, respectively}. The regions of positive $\Delta\sigma_{D}^{\prime}$ for $\mu$ values equal to 0.3 and 0.4 show better spatial correspondence with the seismicity of the CSZ. However, the model with $\mu = 0.4$ predicts positive Coulomb stress changes along the entire length of the rift faults, which is contrary to the limited spatial extent of the observed seismicity. \add[OF]{Also, about 66\% of the hypocenters in the model with a $\mu$ of 0.3 fall within the region of $\Delta\sigma_{D}^{\prime} \ge$ 0.5\%, compared to a value of 35\% when $\mu$ is 0.4. However, when we consider the region with $\Delta\sigma_{D}^{\prime} \ge$ 0\%, the percentage increases to 82\% when $\mu$ is 0.4 compared to 73\% in the model with $\mu$ equal 0.3.} $\Delta\sigma_{D}^{\prime}$ approaches zero when $\mu$ values are 0.5 and 0.6, for which the faults are locked. \add[OF]{The percentage of hypocenters in the region with $\Delta\sigma_{D}^{\prime} \ge$ 0.5\% reduces to 1.4\% and 0\% when $\mu$ values are 0.5 and 0.6, respectively.}

\begin{figure}[ht]
\centering
%\includegraphics[width=30pc]{Figures/SD70R25V.png}
\includegraphics[width=30pc]{Figures/SD70R25V_2.png}
\caption{Effect of friction coefficient on the $\Delta\sigma_{D}^{\prime}$ and $\phi_{\max}$ in the SD70R25V model at 10 km depth. Note the change in scale for $\mu$ greater than 0.3. Earthquakes within 2 km of each depth slice are represented by black dots in \ref{fig:eff_of_mu}A. \add[OF]{The outline of the impact structure is represented as a black circle.} Each figure is 70 $\times$ 70 km.} 
\label{fig:eff_of_mu}
\end{figure}

@@

Our present models, however, cannot explain the differences in earthquake activity between the \change[OF]{crater}{impact structure} and the surrounding crust in the CSZ. `b' values of the Gutenberg-Richter relationship for the regions within and outside the \change[OF]{crater}{impact structure} are 0.93 and 0.74, respectively~\citep{Yu_2016}, indicating a higher frequency of small magnitude earthquakes within the \change[OF]{crater}{impact structure}. Also, \change[OF]{the large-magnitude}{most M4+} earthquakes in the CSZ \change[OF]{occurred}{appear to occur} in the northeast region outside the \change[OF]{crater}{impact structure}~\citep[e.g.,][]{lamontagne1999,Powell_2017,Baird_2010}. \add[OF]{Reexamination of historical earthquakes in the Charlevoix area by}~\citet{Stevens1980} \add[OF]{shows that some M4+ earthquakes in the period 1924-1978 also occurred at the SE end of the impact structure.} Our models do not show significantly higher values of $\Delta\sigma_{D}^\prime$ in the seismogenic zone outside the \change[OF]{crater}{impact structure} relative to the regions within the impact \change[OF]{crater}{structure} (Fig. \ref{fig:SD70R25}C). A possible explanation is that faults created in the highly fractured \change[OF]{crater}{impact structure} region~\citep{RONDOT_2000} might have lower friction coefficients than the faults outside the \change[OF]{crater}{impact structure}. Stress build-up on those weaker \change[OF]{crater}{impact structure} faults can be quicker and smaller in magnitude than on faults outside the \change[OF]{crater}{impact structure} under the same tectonic loading, resulting in more numerous, smaller magnitude earthquakes within versus outside the \change[OF]{crater}{impact structure}.

@@

\subsection{Frictional strengths of the rift faults}
The surface expression of the rift faults used in this study correspond to the Gouffre Northwest Fault, Saint-Laurent fault and Charlevoix Fault (Fig. 1). We did not consider the South Shore fault given its aseismic nature and \change[OF]{its location at the extreme eastern edge of the crater}{it appears to mark a boundary to the active zone}~\citep{lamontagne1999}. Our models show that a $\mu$ of 0.3 and cohesion values of 0 MPa create $\Delta\sigma_{D}^\prime$ maps that spatially correlate with the seismicity of CSZ and the observed stress rotations (Fig. \ref{fig:eff_of_mu}). \add[OF]{We ran a series of models in which the rift faults dip at 70$^{\circ}$ and the friction coefficient ($\mu$) is fixed to be 0.3 while cohesion ($C$) is one of 0, 3, 5, 10, 20 and 30 MPa (Fig. S6). Cohesion values less than 5 MPa could explain the spatial distribution of observed seismicity in the CSZ.} \add[OF]{A $\mu$ of 0.4 could also be suitable for the rift faults as the $\Delta\sigma_{D}^\prime$ maps also spatially correlate with the seismicity of the CSZ, and some earthquakes also occurred at the SW region outside the impact structure (Fig. $\ref{figone}$).} %However, previous works in the CSZ agreed with a finite extent of the seismic zone} ~\citep[e.g.,][]{Stevens1980}. 
\change[OF]{This}{Our fault model} is different from the work of \citet{Baird_2010} who modeled the Gouffre Northwest Fault, Saint-Laurent fault and South Shore Fault (Fig. \ref{figone}).

@@

\subsection{Velocity model of the CSZ}
Our models show that a 1D velocity model of the CSZ \citep{lamontagne1999} (Fig. \ref{figthree}) with embedded \change[OF]{crater}{impact structure} and rift faults can explain the observed seismicity and more than 50$\%$ of the stress rotations (Fig. \ref{fig:eff_of_vel}A). Using the velocity model of the \change[OF]{Sanguenay}{Saguenay} region does not improve this result (Fig. \ref{fig:SD70R25}). Similarly, $\Delta\sigma_{D}^\prime$ in the TD70 model did not spatially fit the recorded earthquakes in the upper 10 km in the CSZ because the distinct and lower-strength damaged zone is not apparent in the 3D tomography \citep{Powell_2017} (Fig. \ref{fig:eff_of_vel}B). \citet{lamontagne1999} \change[OF]{Further investigation is needed to determine why the supposedly better velocity models fail}{constructed a 1-D velocity model for the Canadian Shield region that does not include lateral variations of} velocity model to explain the seismicity and stress orientations.

@@

\item The combined effects of the impact \change[OF]{crater}{impact structure} and the rift faults are required in the CSZ to explain the observed distribution of seismicity and more than 50$\%$ of the observed clockwise rotations of the $\phi_{\max}$ in the CSZ \add[OF]{with respect to the regional maximum compressive stress direction}.
